### 2018-11-01 paper/paper.tex
paper: Address feedback from Jarno

\documentclass[10pt,sigconf,authorversion]{lpc}
\usepackage{balance}
\usepackage{courier}
\usepackage{helvet}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{times}
\usepackage{xcolor}

\pdfinfo{
/Title (Building socket-aware BPF programs)
/Author (Joe Stringer)}

\title{Building socket-aware BPF programs}
\author{Joe Stringer}
\affiliation{%
      \institution{Cilium.io}}
\email{joe@cilium.io}

\acmDOI{}
\setcounter{secnumdepth}{0}
\setcopyright{none}

\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}
\newcommand\linuxversion{{v4.20}}

\begin{filecontents*}{../sk-lookup-api.h}
\import{../sk-lookup-api.h}
\end{filecontents*}
\begin{filecontents*}{sk-lookup-structures.h}
\import{sk-lookup-structures.h}
\end{filecontents*}

\begin{document}

\begin{abstract}

    Over the past several years, BPF has steadily become more powerful in
    multiple ways: Through building more intelligence into the verifier which
    allows more complex programs to be loaded, and through extension of the API
    such as by adding new map types and new native BPF function calls. While
    BPF has its roots in applying filters at the socket layer, the ability to
    introspect the sockets relating to traffic being filtered has been limited.

    To build socket introspection into a BPF helper, the verifier needs the
    ability to track the safety of the calls, including appropriate reference
    counting upon the underlying socket. This paper describes extensions to the
    verifier to perform tracking of references in a BPF program. This allows
    BPF developers to extend the UAPI with functions that allocate and release
    resources within the execution lifetime of a BPF program, and the verifier
    will validate that the resources are released exactly once prior to program
    completion.

    Using this new reference tracking ability in the verifier, we add socket
    lookup and release function calls to the BPF API, allowing BPF programs to
    safely find a socket and build logic upon the presence or attributes of a
    socket. This can be used to conditionally redirect traffic based on the
    presence of a listening application, or to implement stateful firewalling
    primitives to understand whether traffic for this connection has been seen
    before. With this new functionality, BPF programs can integrate more
    closely with the networking stack's understanding of the traffic transiting
    the kernel.

\end{abstract}

\maketitle

\section{Keywords}

BPF, firewalls, Linux, networks, packet processing, sockets

\section{Introduction}

The Berkeley Packet Filter~\cite{cbpf}, introduced in 1992, provided a model
for the implementation of flexible packet filtering by allowing runtime
modification of kernel logic from userspace applications. BPF quickly became
the de facto technology for monitoring the traffic flow within POSIX systems.
Originally, this was restricted to applying filtering upon sockets attached to
devices and not to implement filtering capabilities directly for traffic
flowing through a system. Subsequent work investigated using BPF as a basis for
packet filtering on devices~\cite{ipfw}, however this design does not appear to
have been integrated into major operating systems until the introduction of
extended BPF into Linux~\cite{lwnebpf}.

In recent years, multiple efforts have been made to build new filtering
implementations based upon
BPF~\cite{cilium,ovs-ebpf,nsx-harmony,droplet,xdp-ddos,alfwrapper}. Many of
these implementations aim to provide the functionality of a stateful
firewall~\cite{stateful-fw} - that is, network policies may be defined to allow
traffic from one network endpoint to talk to another network endpoint, and to
implicitly allow responses to pass back through the firewall. Many
implementations over the past decade have developed this functionality in a
manner that is agnostic to the locality of the endpoints. This paper revisits
the possibilities for stateful firewalling when the endpoint socket is
co-located with the kernel providing firewalling functionality, as is the case
with container workloads. This effectively combines the ideas of a BPF-based
firewall~\cite{ipfw} with the model of a stateful firewall~\cite{stateful-fw},
operating with an understanding of socket state~\cite{dfw}.

This paper is laid out as follows. Firstly, we discuss the constraints that the
BPF verifier places upon BPF programs and how those constraints translate into
requirements on the implementation of socket lookup helpers in the Linux BPF
API. The implementation of reference tracking functionality is presented.
Secondly, we describe extensions to the BPF API to provide socket introspection
capabilities, taking into account flexibility and performance. Thirdly, we
present some use cases for the new functionality, including more detail on the
stateful firewall case and an additional usage in handling management traffic
in forwarding devices. Finally, we conclude and assess the position of this
work.

\section{Extending the Verifier}

%To validate access around sockets, the verifier must have knowledge of the
%socket type to validate that access is within bounds and is safe with respect
%to read or write.

The BPF verifier must ensure that all BPF programs that are loaded into the
kernel are safe to execute. This necessitates verification of every path
through the program to ensure constraint such as that the program will halt,
and that any dereferencing of pointers is done in a safe and controlled manner.
For accessing socket pointers, this requires validation that memory accesses
are within the bounds of the memory allocated for the structure, and that no
modifications are made to the socket fields that would be cause the core socket
handling logic to get into a bad state. This is provided through extension of
existing functionality which performs bounds verification and offset conversion
for programs operating on CGroups hooks. Additionally, the underlying memory
must remain associated with the socket for the duration of accesses from the
BPF program, which may be guaranteed by taking a reference on sockets for the
duration of their use. The following sections explain how this is implemented
through the introduction of a new pointer type in the verifier, and through the
extension of the verifier to recognise assembly instructions that represent
reference acquisition and release.

\subsection{Pointers to Sockets}

Earlier work~\cite{bpf-sock} introduced a \verb+bpf_sock+ structure to the BPF
API and implemented bounds and access type checking, along with offset
rewriting for converting the access of socket attributes from BPF instructions
into equivalent accesses of the underlying kernel types. This work was
introduced with new BPF hook points which provide the socket structure as the
context of the BPF program.

To allow BPF programs to retrieve and access a socket pointer, the verifier
must be made aware of when a register contains a pointer of this type, and also
how to validate pointer access. The existing verifier logic handles context
pointers in a generic manner via \verb+bpf_verifier_ops+. The core of the
verifier is not aware of specifice pointer types that are used as the context
for the program, only how to verify the pointer access for these types.
However, the packet event hook points targeted in this paper already contain a
context pointer of a type that is different from \verb+bpf_sock+, so some
additional logic was required in the verifier to understand this pointer type.
The patch series associated with this paper introduced a new pointer type
specific to sockets, and linked its validation directly to the aforementioned
socket verification functions without using the verifier operations
abstraction. Write operations into the socket are rejected in the current
implementation. If variations in access and access type into the socket is
desired in future based upon the hook point, then it may make sense to
introduce a similar abstraction for the socket type.

With the verifier now aware of socket pointer types, this provides a foundation
to associate socket pointers with references to kernel resources.

\subsection{Reference tracking}

Sockets in Linux are reference-counted to track usage and handle memory
release. In general to ensure the safety of socket for the duration of access
in BPF programs, this reference count needs to be incremented when sockets are
referenced from BPF programs. Two options were considered for how to guarantee
this: Implicit reference tracking and explicit reference tracking.

%To ensure that the socket remains safe to access for a period of BPF program
%execution, the BPF program must also take a reference on the socket. This also
%implies that such references must be subsequently released to prevent leaking
%kernel resources.

\subsubsection{Implicit reference tracking.}

This calls for the BPF infrastructure to handle reference tracking, and to hide
this detail from the BPF API. Whenever a socket lookup function is called, a
reference is taken and the socket is added to a reference list. At the end of
the BPF program execution, the core code could walk this reference list and
release each reference. However, collecting references and releasing them at
the end of the BPF program invocation has the unfortunate overhead that even
the execution of BPF programs that do not make use of the new socket lookup
helpers would need to pay the cost of checking the list of references to free.
When dealing with sufficiently demanding use cases such as those which use
eXpress Data Path (XDP)~\cite{xdp}, even the few instructions required to
implement such a check may have a measurable impact on performance.

\subsubsection{Explicit reference tracking.}

This calls for reference acquisition and release semantics to be built into the
API. If references must be taken, then requiring BPF program writers to
explicitly handle these references ensures that the program writer must
understand the potential impact this may have on the operation of the program
(including atomic operation cost), and it also ensures that the cost is only
borne by programs that make use of this feature.

\subsubsection{Implementation.}

Based upon the tradeoffs described above, explicit reference tracking was
chosen. From an implementation perspective, this requires tracking pointer
values through each conditional path in the program, and rejecting the load of
programs that fail to "balance" resource acquisition (lookup) with release.
This implementation works as follows.

Calls to helper functions which acquire references to resources are annotated
in the verifier to associate the acquisition and release functions. When the
instruction that acquires a resource is processed, a resource identifier is
allocated for this resource. This identifier is kept in a list in the verifier
state, and it is also associated with the register which receives the resulting
resource pointer. When processing a corresponding release helper call, if a
parameter to the function contains a register that is associated with a
resource identifier, then the identifier is removed from the verifier state. If
any paths in the program reach the final (\verb+BPF_EXIT+) instruction and the
verifier state contains any resource identifier association, then the program
is considered unsafe as it has leaked the resource.

Some instruction calls are restricted while holding a resource reference to
avoid leaking references, for example the \verb+bpf_tail_call+ helper function,
which would otherwise leak the pointer reference to a subsequent BPF program.

\subsection{Runtime protection}

The verifier is tasked with ensuring that all BPF program execution flow is
safe to the extent where it interacts with the BPF API. In general BPF
programs---once verified and triggered by event hooks---will execute from the
first instruction through the final instruction for a particular path through
the program. Therefore, references to those resources should not leak if the
program has been verified.

One exception to this is when classic BPF \verb+LD_ABS+ or \verb+LD_IND+
instructions are used in the BPF program. These instructions were previously
used to provide direct packet access, with the semantics that if the access
offset exceeds the length of the packet, the BPF program would be terminated
prior to reaching the final state. Without mitigation, this could lead to
leaking of references at runtime, even if the instructions appear to balance
acquisition and release at verification time. Newer BPF APIs provide better
alternatives to these instructions for packet access, so we disallow the use of
these instructions while holding a socket reference.

\section{Extending the BPF API}

To provide an API that implements explicit reference tracking, at least two
functions are needed to handle referencing of sockets: A socket lookup function
and a socket release function. This section describes the API considerations
for each of these.

\subsection{API definition}

The following considerations were made for implementing the lookup helpers. The
consideration of each of these items reflects the definition of the lookup
helpers in Listing~\ref{lst:helper-api} and the structure definitions in
Listing~\ref{lst:helper-struct}.

\begin{itemize}
    \item The Linux stack may contain multiple network namespaces, where the
          BPF program may operate attached to a device in one network namespace
          while the desired socket may exist in another network namespace.
    \item BPF programmers may wish to discover sockets that are not directly
          related to the BPF program context (packet, socket, etc).
    \item Current use cases are focused around UDP and TCP sockets, however
          there may be a desire to extend this in future.
\end{itemize}

\lstinputlisting[float=*b,%
  caption={BPF API helper functions for socket lookup},%
  language=c,%
  label=lst:helper-api,%
  breaklines=true]{../sk-lookup-api.h}

\lstinputlisting[float=*b,%
  caption={BPF API structures for socket lookup (Linux~\linuxversion)},%
  language=c,%
  label=lst:helper-struct,%
  breaklines=true]{sk-lookup-structures.h}

\subsubsection{Namespacing.}

Linux network devices are associated with a particular network namespace which
provides logical network stack separation~\cite{netns, netns-netdevice}.
Sockets are also inherently associated with an application that operates in a
particular network namespace~\cite{netns-sock}. The simplest path for handling
namespacing from socket lookup handlers operating from a network device would
be to allow BPF programs operating on a device within a network namespace to
only find sockets in the same network namespace. By passing the BPF program
context into the socket lookup helper, the implementation can derive the source
network namespace from this context. Beyond this, with the growing popularity
of containers it may be desirable for a network orchestration tool to make use
of socket properties inside container network namespaces to influence decisions
made on packets outside the container---for instance, on packets being received
into a node prior to passing the packet into the network namespace. By allowing
the netns ID to be specified, this use case can be supported as well.

\subsubsection{Tuple definition.}

A common use case for the socket lookup functions would call them from a packet
context, where the \verb+__sk_buff+ has the full 5-tuple information available
and easily accessible. One could imagine a simpler helper function that allows
the BPF program writer to provide the \verb+__sk_buff+ to the helper function,
then the helper function would look up the socket based upon the packet
metadata. This would however limit the potential uses for such a helper. Two
cases that would be more limited with this model are usage from XDP (which does
not inherently parse packets or collect packet metadata into the context), or
if the BPF program implements a form of network address translation. From XDP,
if the raw \verb+xdp_md+ context is provided to such a helper, then the helper
would need to parse the packet to understand the 5-tuple. This would duplicate
the standard stack paths which are executed after XDP where the \verb+sk_buff+
is built, but this information would be subsequently thrown away upon return of
the helper call. Adding to this, if the program performs network address
translation between the local application and the remote destination, then the
local stack may not contain a socket associated with the tuple that is directly
in the packet. While these cases could be worked around with a simpler helper,
it was deemed more powerful and generic to allow the BPF program writer to
provide the tuple for lookup.

\subsubsection{Extensibility.}

New kernel APIs that have any scope for extension should contain a flags
argument~\cite{syscall-flags}. For the socket lookup helpers, a few ideas had
been considered as a potential future alternatives to the existing behaviour.
The lookup functions follow the standard socket lookup paths in Linux which
have predetermined methods for selecting a socket when the application uses
\verb+SO_REUSEPORT+. Some subsequent discussion on the mailinglist proposed
allowing BPF program writers to influence the socket selection
mechanism~\cite{netdev-sk-select}.

\subsubsection{Result.}

Another aspect of extensibility is the ability to lookup sockets which are not
TCP or UDP. The initial RFC of this patch series proposed a single lookup
function which would choose the Layer~4 protocol based upon a field in the
tuple~\cite{sk-lookup-rfc}, however this would make it more difficult for BPF
program writers to detect the support for different protocol types at compile
time. When a socket type is unsupported, the initial implementation would
return \verb+NULL+ at runtime, implying that there is no such socket. There is
not a significant number of supported Layer~4 socket types in the Linux stack
today, and the number is not expected to increase drastically, so it was
considered simpler and easier to split out each Layer~4 socket lookup function
into an independent helper call, so that the result could simply return a
socket or \verb+NULL+ and not need to encode a representation of other error
conditions. This simplifies not only the API, but also the verifier
implementation that must understand the return type of the helper functions.

\subsection{Optimizations}

Multiple optimizations were proposed during development of this feature to
reduce the runtime overhead of using the socket lookup helpers. Two such
optimizations are described below: Skipping reference counting when unnecessary
to save on atomic instructions, and allowing the use of direct packet pointers
for the tuple.

\subsubsection{Avoid reference counting when unnecessary.}

For some socket types, such as UDP sockets, or TCP listen sockets, memory
access safety can be achieved with minimal implementation: The destruction of
such sockets is already governed by standard RCU rules, meaning that while the
RCU lock is held, they can be safely accessed without holding a reference; Once
the RCU grace period is reached, the memory may be freed and references to the
socket are no longer safe to use. For these socket types, since BPF programs
run under the RCU lock, the properties of these sockets can be accessed
directly without taking a reference on the socket. As such, the implementation
of the socket lookup and release functions can avoid the atomic reference count
increment operation.

The API retains explicit reference tracking to ensure consistency of the API
and to allow multiple underlying implementations to handle reference tracking
in a way that is safe for the implementation. For instance, for  TCP sockets
that are not governed by RCU, the networking stack uses reference counting to
manage socket memory instead, so the socket lookup and release function
implementations must take and release references on the socket, respectively.

\subsubsection{Allow lookup using direct packet pointers.}

The tuple structure is defined in such a way that if the IPv4 or IPv6 packet is
immediately followed by the TCP or UDP header without IP options in between,
then a pointer to the packet data at the offset of the Layer~3 addresses may be
passed to the lookup function, allowing the implementation to directly pull the
addresses and ports from the packet buffer rather than requiring the BPF
program to first extract these onto the stack and pass a pointer to the stack
copy of the tuple.

\subsection{Future work}

The implementation introduced in the Linux~\linuxversion~release cycle includes
support for looking up TCP and UDP sockets for IPv4 and IPv6 traffic, using the
\verb+SCHED_CLS+, \verb+SCHED_ACT+ and \verb+SK_SKB+ hook points. A proposed
patchset extends this to allow the same helpers to be used from the \verb+XDP+
hook points~\cite{sk-lookup-xdp}.

%\todo{Associate custom metadata with socket?}

\section{Use cases}

\subsection{Stateful Firewalling}

Stateful firewalling is distinguished from stateless firewalling in that it
provides the ability to associate two directions of a connection together and
apply a filtering policy based upon the direction of the connection. This
commonly involves using a connection tracker to reconstruct the connection
state of each endpoint of the connection without specific knowledge from those
endpoints~\cite{conntrack,ovs-ct}.

In cases where Linux is configured as a firewall between applications that
exist in distinct Linux kernel copies running in the network, this is
intuitive. The information that is needed to perform stateful firewalling is
not present on the Linux instance that needs this information, so to understand
the state of the peers the information must be either shared from the peers
(assuming that the firewall orchestrator has control over the peer endpoints),
or this information must be reconstructed based upon the packets that are seen
on this intermediate Linux instance. However, in many cases this model is used
when filtering traffic where one peer is co-located with the Linux stack that
performs the firewalling, which already contains information about the local
end of the connection. In this use case, the connection tracking table
effectively duplicates knowledge that Linux already creates and maintains.

Using the socket lookup functions described in this paper, in conjunction with
the socket properties including local Layer~3 and Layer~4 information available
in the \verb+bpf_sock+ (Listing~\ref{lst:helper-struct}), BPF programs can
identify the directionality of connections for packets ingressing or egressing
the Linux stack on a network device. Between this directionality information
and a network policy provided by a userspace program, BPF programs are able to
identify locally-sourced or locally-destined traffic and allow or deny traffic
ingressing or egressing the Linux stack.

\subsection{Conditional forwarding of management traffic in forwarding devices}

A common design of devices providing routing and forwarding functionality is to
place high-bandwidth forwarding elements into a system with a low-bandwidth
CPU. In many cases, the goal of the CPU is to only to handle control traffic
for configuring the forwarding path, which is a fraction of the traffic handled
by the actual forwarding path. Existing implementations may direct only traffic
for a specific IP into the management CPU, or could redirect all traffic to the
management CPU and configure DoS filters in XDP so that only legitimate traffic
is forwarded up the stack. Such implementations may not provide the degree of
restriction on locally destined traffic that is desired, or may require
additional co-ordination with the socket layer that requires significant
orchestration.

Using the upcoming XDP support for socket lookup, early in the driver level of
the Linux stack the traffic could be compared with the socket table and only
passed up the stack if the traffic corresponds to a local application. As this
functionality uses the socket table directly, it is always kept in sync with
the current applications without additional orchestration of BPF maps. Traffic
which is not intended to be handled locally can then be handled by additional
BPF logic programmed in the XDP hook.

\section{Conclusion}

This paper describes the contribution of new BPF verifier functionality in the
Linux kernel to track references to kernel resources and uses these to provide
access to socket introspection capabilities from packet handling hooks. The
introduction of reference tracking logic into the verifier provides
useful base infrastructure for supporting acquire and release semantics for
kernel resources which may prove useful for other helpers in future, and the
socket lookup API makes the packet hooks more powerful to support use cases
such as stateful firewalling or conditional redirect based on the sockets that
are open on a system.

%\todo{Associate custom metadata with socket?}

\section{Acknowledgments}

The author would like to acknowledge Alexei Starovoitov, Daniel Borkmann, Jarno
Rajahalme, Martin KaFai Lau, Nitin Hande and Thomas Graf for guidance, review
and testing of the patches, and feedback on this paper.

\bibliographystyle{plainnat}
\bibliography{paper}

\end{document}